Bayesian inference of relative fitness on high-throughput pooled competition assays

### 2023-10-13 doc/paper.tex
prepare paper for submission

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

@@

\usepackage{lineno}
\linenumbers
\usepackage[font={footnotesize, sf}, labelfont=bf, margin=0.05\linewidth]{caption}
\renewcommand{\familydefault}{\sfdefault}
% Bold the 'Figure #' in the caption and separate it from the title/caption
% with a period
% Captions will be left justified
\usepackage[
	aboveskip=1pt,
	labelfont=bf,
	labelsep=period,
	justification=raggedright,
	singlelinecheck=off
]{caption}

@@

% Add author affiliations
\usepackage{authblk}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

% References
% \usepackage[]{biblatex}
\usepackage[
	backend=bibtex,
	style=phys, 
	sorting=none,				% Do not sort bibliography
	url=false, 					% Do not show url in reference
	doi=false, 					% Do not show doi in reference
	isbn=false, 				% Do not show isbn link in reference
	eprint=false, 			% Do not show eprint link in reference
	maxbibnames=9, 			% Include up to 9 names in citation
	firstinits=true,
]{biblatex}
\addbibresource{references.bib}

\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdfauthor={Manuel Razo-Mejia; Madhav Mani; Dmitri Petrov},
  pdfkeywords={bayesian inference, variational inference, relative
fitness, open-source, reproducible research},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

% Title
\title{Bayesian inference of relative fitness on high-throughput pooled competition assays}

% Authors
\author[1,*]{Manuel Razo-Mejia}
\author[2,3]{Madhav Mani}
\author[1,4,5]{Dmitri Petrov}

% Affiliations
\affil[1]{Department of Biology, Stanford University, Stanford, CA, USA}
\affil[2]{Department of Engineering Sciences and Applied Mathematics, Northwestern University, Chicago, IL, USA}
\affil[3]{NSF-Simons Center for Quantitative Biology, Northwestern University, Chicago, IL, USA}
\affil[4]{Stanford Cancer Institute, Stanford University School of Medicine, Stanford, CA, USA}
\affil[5]{Chan Zuckerberg Biohub}
\affil[*]{Correspondence: mrazo@stanford.edu}

\setcounter{Maxaffil}{0}
% Set affiliations in small font
\renewcommand\Affilfont{\itshape\small}

\date{}

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, interior hidden, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

% Remove main text from the table of contents by specifying not to include
% any section or subsection
\addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}

% Define reference segment for main text
\begin{refsegment}
% Generate filter to not include references from main text in the
% supplemental references
\defbibfilter{notother}{not segment=\therefsegment}

\maketitle

\begin{abstract}
    The tracking of lineage frequencies via DNA barcode sequencing enables the
  quantification of microbial fitness. However, experimental noise coming from
  biotic and abiotic sources complicates the computation of a reliable
  inference. We present a Bayesian pipeline to infer relative microbial fitness
  from high-throughput lineage tracking assays. Our model accounts for multiple
  sources of noise and propagates uncertainties throughout all parameters in a
  systematic way. Furthermore, using modern variational inference methods based
  on automatic differentiation, we are able to scale the inference to a large
  number of unique barcodes. We extend this core model to analyze
  multi-environment assays, replicate experiments, and barcodes linked to
  genotypes. On simulations, our method recovers known parameters within
  posterior credible intervals. This work provides a generalizable Bayesian
  framework to analyze lineage tracking experiments. The accompanying
  open-source software library enables the adoption of principled statistical
  methods in experimental evolution.
\end{abstract}

@@

% Print main text references
\printbibliography[segment=\therefsegment]
% Close reference segment
\end{refsegment}

\clearpage

% SUPPLEMENTAL MATERIAL

% Indicate that now all sections and subsections should be included in the
% table of contents so that only the SI is included.
\addtocontents{toc}{\protect\setcounter{tocdepth}{3}}

% Define reference section for the supplemental material
\begin{refsegment}
% Set equation, table and figure counters to begin with "S"
\beginsupplement

@@

\tableofcontents

@@

% Print supplemental references changing the title
\printbibliography[title={Supplemental References},
segment=\therefsegment, filter=notother]
\end{refsegment}

% \printbibliography

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{scrartcl}

@@

% Change font size
\usepackage[font={footnotesize, sf}, labelfont=bf, margin=0.05\linewidth]{caption}
\renewcommand{\familydefault}{\sfdefault}
% Bold the 'Figure #' in the caption and separate it from the title/caption
% with a period
% Captions will be left justified
\usepackage[
	aboveskip=1pt,
	labelfont=bf,
	labelsep=period,
	justification=raggedright,
	singlelinecheck=off
]{caption}

@@

% Add author affiliations
\usepackage{authblk}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[
    style=phys,
    sorting=none,				% Do not sort bibliography
	url=false, 					% Do not show url in reference
	doi=false, 					% Do not show doi in reference
	isbn=false, 				% Do not show isbn link in reference
	eprint=false, 			    % Do not show eprint link in reference
	maxbibnames=9, 			    % Include up to 9 names in citation
	firstinits=true,
]{biblatex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addbibresource{references.bib}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Bayesian inference of relative fitness on high-throughput pooled competition assays},
  pdfauthor={Manuel Razo-Mejia; Madhav Mani; Dmitri Petrov},
  pdfkeywords={bayesian inference, variational inference, relative
fitness, open-source, reproducible research},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Add title
\title{Bayesian inference of relative fitness on high-throughput pooled
competition assays}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Add list of authors

% Loop through authors
\author[1,*]{Manuel Razo-Mejia}
\author[3,4]{Madhav Mani}
\author[1,2,5]{Dmitri Petrov}

% Loop through affiliations
\affil[1]{Department of Biology, Stanford University}
\affil[2]{Stanford Cancer Institute, Stanford University School of
Medicine}
\affil[3]{NSF-Simons Center for Quantitative Biology, Northwestern
University}
\affil[4]{Department of Molecular Biosciences, Northwestern University}
\affil[5]{Chan Zuckerberg Biohub}

% Add corresponding author
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Set affiliations in small font
\renewcommand\Affilfont{\itshape\small}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Remove date
\date{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle
\begin{abstract}
The tracking of lineage frequencies via DNA barcode sequencing enables
the quantification of microbial fitness. However, experimental noise
coming from biotic and abiotic sources complicates the computation of a
reliable inference. We present a Bayesian pipeline to infer relative
microbial fitness from high-throughput lineage tracking assays. Our
model accounts for multiple sources of noise and propagates
uncertainties throughout all parameters in a systematic way.
Furthermore, using modern variational inference methods based on
automatic differentiation, we are able to scale the inference to a large
number of unique barcodes. We extend this core model to analyze
multi-environment assays, replicate experiments, and barcodes linked to
genotypes. On simulations, our method recovers known parameters within
posterior credible intervals. This work provides a generalizable
Bayesian framework to analyze lineage tracking experiments. The
accompanying open-source software library enables the adoption of
principled statistical methods in experimental evolution.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden]}{\end{tcolorbox}}\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Remove main text from the table of contents by specifying not to include
% any section or subsection
\addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}

% Define reference segment for main text
\begin{refsegment}
% Generate filter to not include references from main text in the
% supplemental references
\defbibfilter{notother}{not segment=\therefsegment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Print main text references
\printbibliography[segment=\therefsegment]
% Close reference segment
\end{refsegment}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SUPPLEMENTAL MATERIAL

% Indicate that all sections and subsections should be included in the
% table of contents so that only the SI is included.
\addtocontents{toc}{\protect\setcounter{tocdepth}{3}}

% Define the reference section for the supplemental material
\begin{refsegment}
% Set equation, table, and figure counters to begin with "S"
\beginsupplement
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Add table of contents
\tableofcontents
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% HERE IS WHERE THE SI TEXT GOES
% Print supplemental references changing the title
\printbibliography[title={Supplemental References},
segment=\therefsegment, filter=notother]
\end{refsegment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
